Model Adaptivity for Goal-Oriented Inference using Adjoints

### 2017-05-30 paper/submitted/reviews/reviewer_response.tex

\documentclass[12pt, letterpaper]{article}
\usepackage{amsfonts}
\usepackage{amsmath,amsthm,amssymb,amsfonts,amsbsy,latexsym}
\usepackage{color}
\usepackage{changepage}
\usepackage{parskip}
\usepackage[margin=0.75in]{geometry}

\newcommand{\answer}[1]{\begin{adjustwidth}{1cm}{}{\color{blue}#1}\end{adjustwidth}}
\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\blue}[1]{{\color{blue}#1}}
\newcommand{\green}[1]{{\color{green}#1}}

\newcommand{\notdone}{{\color{red}{Changes not yet made.}}}
\newcommand{\done}{{\color{green}{Changes made.}}}

\begin{document}

\section*{Response to Referee Comments}

We thank the reviewers for their detailed and insightful comments and helping us strengthen the manuscript. Our point-by-point responses to the critique are outlined below, with reviewer comments in black and \blue{responses in blue}. Reviewers are ordered by the order in which their comments were received.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Reviewer 4}

The paper is basically an application of [7] to modeling error demonstrated with rather transparent problems in 2D flow. The authors ignore the first papers on Goal-Oriented Estimation and modeling error laid out in the papers of Oden and Vemaganti in JCP in 2000, and extended by these authors in CMAME in 2001. The relationship between the current work and these papers should be clearly explained and cited. These revisions and re-review would be advised.

\answer{We apologize for not referencing the aforementioned papers; we had not encountered them in our literature review and did not intend to imply their irrelevance. Citations of these papers and explanation of their relevance to our work has been added in \red{Section x.x.x}}
\notdone

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Reviewer 1}

On page 8 (before equation 10), the sentence “Note that, ..” references equation 11 before it is
presented. It may be better to move this comment to immediately follow the description of equation
11.

\answer{blah}
\notdone

I found the motivation in the beginning of the manuscript somewhat lacking. The description of use
cases, and conditions for which the algorithm is suitable, from the last paragraph on page 9 were very
helpful. I believe the presentation would be strengthened by mentioning some of this earlier in the
introduction.

\answer{}
\notdone

Although it is represented graphically in Figure 8, $\Omega_I$ for the final example is not defined in the text.

\answer{}
\notdone

At the beginning of the second paragraph of Section 4.3.2 (pg. 25) I think the reference to “Section
4.3.2” should be Table 1.

\answer{}
\notdone

The sub-figures of Figure 9 could be larger. It is difficult to verify the claim in the text that “. . . the
domain is refined completely in the x 3 direction first around the QoI region. . . ”. Even going as far as
to include 2-D slices (as in Figure 8) may be warranted.

\answer{}
\notdone

The second to last paragraph on page 28 starts “Section 4.3.2 shows the average QoI. . . ”. Again, I
believe this is a reference to a table, Table 2 in this case.

\answer{}
\notdone

I find the potential to apply this approach to surrogate design, as discussed in the final paragraph, very
interesting. I believe this is a good case for overcoming the cumulative cost of performing multiple
adaptive steps where the overall cost is comparable to evaluating the high-fidelity inverse problem.
If the final adaptive step was significantly cheaper to evaluate (say 2X as in the last example) the
cumulative cost of the adaptive algorithm would be amortized across a large number of posterior
samples with the final mixed-fidelity model.

\answer{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Reviewer 2}

The main results in the paper are Proposition 1 and Eq. (14). These results are very close to existing results as presented in [7]. It is not clearly explained in the introduction, and in Section 2, in what sense the current results follow from [7] and in what sense they are new.

\answer{}
\notdone

The main theoretical section, Section 2, is vague, lacks proper structure and needs clarification. It is not clear how the high-fidelity and low-fidelity models are defined. These models should be the starting point of the section, not introduced in some proposition. Some degree of compatibility is assumed, but not explained. The proof of Proposition 1 refers to Proposition 3 in [7] which is worth stating in full in order to understand how it is employed. Given that (5) was stated as a Proposition 1, it is also natural to state the result in (14) as a proposition too.

\answer{}
\notdone

Specifics of the models are introduced in Section 4 (numerical experiments), which is out of place. It would benefit the reader to move the relevant parts to after Section 2, as examples of the abstract framework. It would also help if the propositions would then be re-interpreted in terms of the specific low/high fidelity model setting: In other words, the authors should explain what is actually computed to do the model adaptivity. This is currently very unclear. (It is also necessary to explain the contents of Algorithm 1.)

\answer{}
\notdone

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Reviewer 3}

I’d like to actually see the statement and proof of Proposition 3 from reference 7 included in
the manuscript. It is at the crux of the proof and it’s frustrating to have to go dig up that
detail.

\answer{}
\notdone

I’d like to see the details of the error estimate localization included in the manuscript instead
of delegating to a reference. The given description is quite “wordy” and not very precise.

\answer{}
\notdone

The spirit of the examples is fine, but I’m not following some of the details. They state for
the example in Section 4.1 that f (q) = q, but then the exact underlying field varies in space.

\answer{}
\notdone

To what space does f (q) belong? Based on the example in Section 4.2.1, it looks like it’s
a constant. If the parameters are spatially dependent, what underlying basis is being used
for its representation? Is the mesh refined enough to rule out pollution of the results due to
discretization error?

\answer{}
\notdone

It could be illustrative to see how the parameters q change with the refinement process.

\answer{}
\notdone

Figure 3 is not particularly illustrative, especially on linear scale. Perhaps some tuning of the
refinement parameters? What were the refinement parameters for the examples?

\answer{}
\notdone

\end{document}

### 2017-05-30 paper/submitted/round2_png/conclusions_v4_cmame.tex
stash reviewer comments; start reviewer response doc and minor edits

We have presented an error estimator that can be used to adaptively create a mixed-fidelity model with which to solve a goal-oriented inverse problem, so as to minimize the error in the QoI calculated from the inferred parameters. We applied this method to pairs of low- and high-fidelity models of convection-diffusion-reaction phenomena and were able to obtain QoI estimates with a small relative error without having to solve the inverse problem with the high-fidelity model. In these cases, the localization of the error estimate also indicated regions of the domain that were important to the interaction of the observations and the QoI.

%------------------------------------------------------------%
\section{Future Work} 
%------------------------------------------------------------%

A direction for extension of this work is to the case of the statistical inverse problem. Thus far in this work, we have considered the deterministic inverse problem, as described in \Cref{sec:setup}; we seek to infer the parameter values that optimally fit the observations and the prior beliefs embedded in the regularization. However, we can rarely, if ever, be certain that the inferred values are correct, whether this be due to epistemic uncertainty from a lack of knowledge or aleatoric uncertainty from inherent variability in the physical system, or both \cite{Ober04}. One may attempt to capture the uncertainty in the inferred parameters by representing them as random variables with a probability distribution; inferring the distribution of the parameters given some observations is the statistical inverse problem.

The statistical inverse problem is often solved in a Bayesian framework. Bayes' rule is used to combine a prior distribution, which captures prior beliefs about the parameters, and a likelihood distribution, which captures the likelihood of observations given an instance of the parameter values and a model of noise in the observations, to give a posterior distribution on the parameters. Since there is generally no analytical expression for this posterior distribution, it is usually characterized by samples from the distribution. Sampling methods like the widely-used Markov chain Monte Carlo (MCMC) method require many evaluations of the forward model, and since the number of samples needed grows exponentially with the dimension of the parameter space, this problem becomes intractable for large parameter spaces. In engineering contexts, it is still usually the case, however, that we are ultimately interested in a low-dimensional QoI, and it is the uncertainty in this low-dimensional quantity that we wish to capture; this low-dimensional distribution is referred to as the predictive posterior.

One way we could potentially apply this work to the statistical inverse problem is by reducing the parameter space that needs to be sampled. Such a direction is suggested by the results presented in \Cref{sec:diffvcdr3D}, where the mixed-fidelity model had significantly fewer degrees of freedom in its parameter field than the high-fidelity model, and thus a smaller parameter space. In the case of a linear model and observation operator with a Gaussian prior and additive Gaussian noise, there are parallels between the objective function of the deterministic inverse problem with Tikhonov regularization and the mode of the posterior distribution. In \cite{Martetal12}, a method is described for creating proposal distributions, drawing from these parallels; both the linear Gaussian and nonlinear cases are addressed. Similarly, these parallels could potentially be drawn upon to extend this work to the creation of an alternative statistical inverse problem that, by utilizing a mixed-fidelity model with fewer degrees of freedom in its parameter field, requires exploration of a small parameter space with minimal compromise in the predictive posterior.

Another potential approach would be to extend our method to the creation of mixed-fidelity models that are used as surrogates; these surrogate models can be evaluated in place of the high-fidelity model, thus decoupling the number of expensive forward evaluations of the high-fidelity model needed from the number of posterior parameter distribution samples that is desired \cite{Con14}. The samples obtained using such a surrogate might sacrifice accuracy in representing the posterior parameter distribution for accuracy in representing the predictive posterior distribution.

%other directions for extensions that I can't seem to fit in:
%mixing models in time as well? (different mixes of models at different time steps?)
%superadj on intermediate mesh; preliminary results suggest that error estimates remain reasonable and generation of mixed-fidelity models can continue despite the additional approximations to the supplementary adjoint

%\item \Cref{alg:refSeries} is also ammenable to an offline-online decomposition, analagous to that proposed in \cite{LiebWill13}. In the case where both the low-fidelity and high-fidelity inverse problems are linear in the data, and the QoI is linear in the state and parameters, one may first compute and store the supplementary adjoint $\Lambda_0$ for the low-fidelity model. As new data is received, one can compute the $\Psi_{LF}$ for this new data; evaluating \cref{eq:finErrExp} with the stored $\Lambda_0$ and the new $\Psi_{LF}$ gives an exact estimate of the error in the QoI, and thus effectively the high-fidelity QoI (NOT TRUE the primary and aux vars appear in the rhs...). When these linearities do not all hold, one cannot obtain an exact error estimate. The offline phase would consist of adaptively creating a mixed-fidelity model with an appropriate error tolerance given some expected observations $d^*$, and storing its suppementary adjoint $\Lambda_{MF}^*$. As new data is received, one can compute the $\Psi_{MF}$ for the mixed-fidelity model and the new data; evaluating \cref{eq:finErrExp} with the stored $\Lambda_{MF}^*$ and the new $\Psi_{MF}$ gives an estimate of the error in the QoI, and thus an effective estimate of the high-fidelity QoI.

>>>>

We adaptively create mixed-fidelity models to solve goal-oriented inverse problems. The paper develops an error estimator that drives the adaptation, so as to minimize the error in the QoI calculated from the inferred parameters. We applied this method to pairs of low- and high-fidelity models of convection-diffusion-reaction phenomena. The results showed QoI estimates with a small relative error even when the high-fidelity model was used only in a small portion of the domain. In these cases, the localization of the error estimate also indicated regions of the domain that were important to the interactions between the observations and the QoI. A direction for extension of this work is to the case of the statistical inverse problem.  One way we could potentially apply this work to the statistical inverse problem is by reducing the parameter space that needs to be sampled. Such a direction is suggested by the results presented in \Cref{sec:diffvcdr3D}, where the mixed-fidelity model had significantly fewer degrees of freedom in its parameter field than the high-fidelity model, and thus a smaller parameter space. One could explore creation of an alternative statistical inverse problem that, by utilizing a mixed-fidelity model with fewer degrees of freedom in its parameter field, requires exploration of a small parameter space with minimal compromise in the predictive posterior.

Another potential approach would be to extend our method to the creation of mixed-fidelity models that are used as surrogates; these surrogate models can be evaluated in place of the high-fidelity model, thus decoupling the number of expensive forward evaluations of the high-fidelity model needed from the number of posterior parameter distribution samples that is desired \cite{Con14}.

### 2017-05-30 paper/submitted/round2_png/intro_v4_cmame.tex
stash reviewer comments; start reviewer response doc and minor edits

Goal-oriented approaches prioritize accuracy in the QoI over accuracy in the states and/or parameters; one aims to reduce the cost of solving the problem (whether forward or inverse) while maintaining an acceptably low error tolerance for the QoI. In the context of the forward problem, goal-oriented methods have been developed for mesh refinement \cite{BecRann01,PrudOden99,VendDarm00,Yano12} and for model adaptivity~\cite{BraackErn03,OdenPrudetal06}. These methods derive an adjoint-based a posteriori error estimate for an output functional, and this estimate is used to guide adaptive mesh and/or model refinement.

The inverse problem setting has also seen some work on goal-oriented methods. It has been recognized that for the goal-oriented inverse problem, fully resolving the parameters is often unnecessary to accurately compute the QoI. For example, for a discretized linear inverse problem, one can find a low-dimensional subspace of the parameter space that is both informed by observations and informative to the QoI~\cite{LiebWill13}. This subspace provides a low-dimensional map from the observations directly to the QoI, sacrificing accuracy in the inferred parameters for accuracy in the QoI that is computed from them. For the linear Gaussian Bayesian inverse problem, one can calculate an optimal approximation to the predictive posterior of the QoI without fully calculating the posterior distribution of the parameters~\cite{Span16}. Mesh refinement in the goal-oriented inverse problem is addressed in~\cite{BecVex05}; that work derives an a posteriori estimate of the error in the QoI caused by discretizing the infinite-dimensional inverse problem, and uses this error estimate to adaptively refine the mesh. The idea of goal-oriented model adaptivity for an inverse problem is introduced in~\cite{OdenPrudetal10}, which outlines a possible extension of the adjoint-based adaptivity method of~\cite{OdenPrudetal06}.

In this paper, we target model adaptivity in the solution of goal-oriented inverse problems. We present a robust, adaptive framework that identifies the subregions of the domain where high-fidelity representation of physics and parameters is important to achieving a desired accuracy in the QoI. To achieve this, we combine ideas from goal-oriented model adaptivity for forward problems and goal-oriented methods for inverse problems. In particular, our method allows one to systematically manage the use of multiple models in the context of the goal-oriented inverse problem, so as to minimize the error in a QoI prediction. Taking the inverse problem with the highest-fidelity model as our reference QoI prediction, we derive a third-order estimate for the QoI error from using a lower-fidelity model. This estimate can be localized, and the error decomposition then used to guide the formation of mixed-fidelity models with which to solve the inverse problem, while minimizing the error in the QoI.

>>>>

Goal-oriented approaches prioritize accuracy in the QoI over accuracy in the states and/or parameters; one aims to reduce the cost of solving the problem (whether forward or inverse) while maintaining an acceptably low error tolerance for the QoI. In the context of the forward problem, goal-oriented methods have been developed for mesh refinement \cite{BecRann01,Park04,PrudOden99,VendDarm00,Yano12} and for model adaptivity~\cite{BraackErn03,OdenPrudetal06}. These methods derive an adjoint-based a posteriori error estimate for an output functional, and this estimate is used to guide adaptive mesh and/or model refinement.

For inverse problems, computational cost has been addressed by employing surrogate and reduced-order models \cite{MarzNajmRahn07, WangZab04, StefSanNav15, Frangosetal10}. For the particular case of a goal-oriented inverse problem, methods exploit the possibility of accurately computing the QoI without fully resolving the parameters. For example, for a discretized linear inverse problem, one can find a low-dimensional subspace of the parameter space that is both informed by observations and informative to the QoI~\cite{LiebWill13}. This subspace provides a low-dimensional map from the observations directly to the QoI, sacrificing accuracy in the inferred parameters for accuracy in the QoI that is computed from them. For the linear Gaussian Bayesian inverse problem, one can calculate an optimal approximation to the predictive posterior of the QoI without fully calculating the posterior distribution of the parameters~\cite{Span16}. Mesh refinement in the goal-oriented inverse problem is addressed in~\cite{BecVex05}; that work derives an a posteriori estimate of the error in the QoI caused by discretizing the infinite-dimensional inverse problem, and uses this error estimate to adaptively refine the mesh. The idea of goal-oriented model adaptivity for an inverse problem is introduced in~\cite{OdenPrudetal10}, which outlines a possible extension of the adjoint-based adaptivity method of~\cite{OdenPrudetal06}.

In this paper, we target model adaptivity in the solution of goal-oriented inverse problems. We present a robust, adaptive framework that identifies the subregions of the domain where high-fidelity representation of physics and parameters is important to achieving a desired accuracy in the QoI. To achieve this, we combine ideas from goal-oriented model adaptivity for forward problems and goal-oriented methods for inverse problems. In particular, our method allows one to systematically manage the use of multiple models in the context of the goal-oriented inverse problem, so as to minimize the error in a QoI prediction. Taking the inverse problem with the highest-fidelity model as our reference QoI prediction, we derive a third-order estimate for the QoI error from using a lower-fidelity model. This estimate can be localized, and the error decomposition then used to guide the formation of mixed-fidelity models with which to solve the inverse problem, while minimizing the error in the QoI. 
%We note that while the approach outlined in~\cite{OdenPrudetal10} could be applied to the same task, one main difference is that we compute additional quantities to obtain a higher-order QoI error estimate.

### 2017-05-30 paper/submitted/round2_png/mathematical_formulation_v4_cmame.tex
stash reviewer comments; start reviewer response doc and minor edits

Extending the property in \Cref{eq:MeqI} to the augmented Lagrangians for the high- and low-fidelity models, we have
%
\begin{multline}
\label{eq:repIwithM}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})\textrm{.}
\end{multline}
%
Applying the output error representation described in Proposition 3 from~\cite{BecVex05} for the difference $\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})$,
\begin{equation}
\label{eq:beckvex}
\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF}) = \frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{R}(e^3)\textrm{.}
\end{equation}
Combining \Cref{eq:repIwithM} and \Cref{eq:beckvex} we obtain
\begin{multline}
\label{eq:preadj}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})+\mathcal{R}(e^3)\textrm{.}
\end{multline}
which completes the proof.
\end{proof}
%

>>>>

Extending the property in \Cref{eq:MeqI} to the augmented Lagrangians for the high- and low-fidelity models, we have
%
\begin{multline}
\label{eq:repIwithM}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})\textrm{.}
\end{multline}
%
Applying the output error representation described in Proposition 3 from~\cite{BecVex05} for the difference $\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})$,
\begin{equation}
\label{eq:beckvex}
\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF}) = \frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{R}(e^3)\textrm{.}
\end{equation}
Combining \Cref{eq:repIwithM} and \Cref{eq:beckvex} we obtain
\begin{multline}
\label{eq:preadj}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})+\mathcal{R}(e^3)\textrm{,}
\end{multline}
which completes the proof.
\end{proof}
%
